---
title: "Function: Define Alpha Diversity Indices"
output: github_document
---
# Purpose of the function
This is a function to calculate the alpha diversity of a given OTU table. It will calculate the Shannon, Simpson, Chao1, and ACE indices.

## Load the needed libraries and convert the .csv  file into a data set
```{r, message=FALSE}
library(tidyverse)
library(vegan)
library(phyloseq)

OTU_data <- read.csv("otutable_grazing.csv")
```

# Function
After converting the .csv file into a data set, the user is able to use this function to obtain the alpha indices.

```{r}
#function to calculate the alpha indices for a given OTU table
calc_alpha_indices <- function(data){
  #make a data frame with just the sample IDs
  sample.ID <- data[, 1]
  samples <- data.frame(sample.ID)
  samples
  
  #data frame with the shannon data and the sample IDs
  shannon.indices <- diversity(data[, -1], index = "shannon", MARGIN = 1, base = exp(1))
  shannon_index <- data.frame(shannon.indices)
  shannon_data <- cbind(samples, shannon_index)
  
  #data frame which includes both shannon and simpson
  simpson.indices <- diversity(data[, -1], index = "simpson", MARGIN = 1, base = exp(1))
  simpson_index <- data.frame(simpson.indices)
  simpson_data <- cbind(shannon_data, simpson_index)
  
  #calculate chao and ACE indices and add them to a final matrix 
  chao.out <- estimateR(data[, -1])
  chao_index <- data.frame(chao.out)
  chao_transpose <- t(chao_index)
  final_matrix <- cbind(simpson_data, chao_transpose)
  
  return(final_matrix)
}
```

Example of running the function and assigning the output to "alpha_matrix"

```{r}
#using the function to create a matrix called alpha_matrix
(alpha_matrix <- calc_alpha_indices(OTU_data))
```

