---
title: "Homework 5"
output: github_document
---
We first want to load in the packages and look at the dataset for diamonds.

```{r, message=FALSE}
library(tidyverse)
library(dplyr)
print(diamonds)
```
# Group Data and show the count
```{r}
diamonds %>%
  group_by(cut) %>%
  count()
```

# Get a random subset of data
```{r}
# this creates a random subset of data and condenses it to 1%
condensed_diamonds <- sample_frac(diamonds, 0.01) 
# view the condensed data
print(condensed_diamonds)
```
# Calculating the average size of 100 largest diamonds
```{r}
# check to make sure we have 100 in each clarity
condensed_diamonds %>%
  group_by(clarity) %>%
  count()
```
Since the condensed subset of data does not have 100 diamonds in each category, we will use the original data.
```{r}
diamonds %>%
  group_by(clarity) %>%
  count()
```
We now have the amount of data needed to proceed. We will now calculate the average size of the 100 largest diamonds in each category. Before calculating the average size, we must first get the data for each clarity.
```{r}
# create new data set for 100 largest diamonds in each category
large_diamonds <- diamonds %>% 
  # we first want to arrange the diamonds data in order from highest to lowest
  arrange(desc(carat)) %>%
  # now that is has been sorted, we can then group them by clarity
  group_by(clarity) %>%
  # we then only want 100 sizes so we slice them out
  slice(1:100)
# print out the new data
large_diamonds 
```

Now that we have the new dataset, we can calculate the average size in each category
```{r}
large_diamonds %>%
  # we need to group them so that we can calculate the average for each category
  group_by(clarity) %>%
  # now that they are grouped, we can calculate the average of each
  summarise(average.size = mean(carat, na.rm = TRUE))
```

# Make two scatter plots

## Plot 1: X vs Y

We first need to make a scatter plot that contains the x vs z dimension of the diamonds data set.
```{r}
ggplot(diamonds) +
  geom_point(mapping = aes(x = x, y = y)) +
  labs(title = "Diamond Dimension: X vs. Y", x = "Length (x)", y = "Height (y)")
```

We can see from the plot that there is a few outliers that we need to get rid of in order to make a better plot.

```{r}
# this filters out the outliers within this dataset
x_y_diamonds <- diamonds %>%
  filter(x > 3 & y < 20)
ggplot(x_y_diamonds) +
  geom_point(mapping = aes(x = x, y = y)) +
  labs(title = "Diamond dimensions: X vs Y", x = "Length (X)", y = "Height (y)")
```

## Plot 2: X vs Z

We now need to make a plot of X vs Z. 
```{r}
ggplot(diamonds) +
  geom_point(mapping = aes(x = x, y = z)) +
  labs(title = "Diamond Dimension: X vs. Z", x = "Length (x)", y = "Width (z)")
```

We run into the same problem as before so we must filter out the outliers
```{r}
x_z_diamonds <- diamonds %>%
  filter(x > 3 & z < 10 & z > 1)
ggplot(x_z_diamonds) +
  geom_point(mapping = aes(x = x, y = z)) +
  labs(title = "Diamond dimensions: X vs Z", x = "Length (X)", y = "Width (z)")
```
